Canopy analysis in R using Forest Tools
========================================
Andrew Plowright: 2018-04-04
### Introduction
The Forest Tools R package offers functions to analyze remotely sensed forest data. Currently, tools to detect dominant treetops and outline tree crowns have been implemented, both of which are applied to a rasterized canopy height model (CHM), which is generally derived from LiDAR or photogrammetric point clouds. A function to summarize the height and count of trees within user-defined geographical areas is also available.
The following vignette provides examples for using these functions.


### Installation
Check that R is up-to-date. This can be done automatically using the installr package. Alternatively, download the latest version directly from the Comprehensive R Archive Network (CRAN).
```{r}
install.packages("installr")
library(installr)
updateR()
```
Download and install the Forest Tools package from CRAN using the install.packages function.
```{r}
install.packages("ForestTools")
```
### Loading sample data
A sample canopy height model (CHM) is included in the Forest Tools package. It represents a small 1.5 hectare swath of forest in the Kootenay Mountains, British Columbia. The following examples use this sample, but if you would rather use your own data, it can be loaded into R using the raster function. A brief section on reading and writing geospatial data in R is included in this document. Otherwise, begin by loading the necessary libraries and the sample CHM using the library and data functions respectively.
```{r}
# Attach the 'ForestTools' and 'raster' libraries
library(ForestTools)
library(raster)

# Load sample canopy height model
data("kootenayCHM")
```
View the CHM using the plot function. The cell values are equal to the canopy’s height above ground.
```{r}
# Remove plot margins (optional)
par(mar = rep(0.5, 4))

# Plot CHM (extra optional arguments remove labels and tick marks from the plot)
plot(kootenayCHM, xlab = "", ylab = "", xaxt='n', yaxt = 'n')
```
### Detecting treetops
Dominant treetops can be detected using vwf. This function implements the variable window filter algorithm developped by Popescu and Wynne (2004). In short, a moving window scans the CHM, and if a given cell is found to be the highest within the window, it is tagged as a treetop. The size of the window itself changes depending on the height of the cell on which it is centered. This is to compensate for varying crown sizes, with tall trees having wide crowns and vice versa.
Therefore, the first step is to define the function that will define the dynamic window size. Essentially, this function should take a CHM cell value (i.e.: the height of the canopy above ground at that location) and return the radius of the search window. Here, we will define a simple linear equation, but any function with a single input and output will work.
lin <- function(x){x * 0.05 + 0.6}
We do not wish for the vwf to tag low-lying underbrush or other spurious treetops, and so we also set a minimum height of 2 m using the minHeight argument. Any cell with a lower value will not be tagged as a treetop.
ttops <- vwf(CHM = kootenayCHM, winFun = lin, minHeight = 2)
We can now plot these treetops on top of the CHM.

```{r}
# Plot CHM
plot(kootenayCHM, xlab = "", ylab = "", xaxt='n', yaxt = 'n')

# Add dominant treetops to the plot
plot(ttops, col = "blue", pch = 20, cex = 0.5, add = TRUE)
```
The ttops object created by vwf in this example contains the spatial coordinates of each detected treetop, as well as two default attributes: height and winRadius. These correspond to the tree’s height above ground and the radius of the moving window where the tree was located. Note that winRadius is not necessarily equivalent to the tree’s crown radius.

```{r}
# Get the mean treetop height
mean(ttops$height)
## [1] 5.404217
```

### Outlining tree crowns
Canopy height models often represent continuous, dense forests, where tree crowns abut against eachother. Outlining discrete crown shapes from this type of forest is often refered to as canopy segmentation, where each crown outline is represented by a segment. Once a set of treetops have been detected from a canopy height model, the mcws function can be used for this purpose.
The mcws function implements the watershed algorithm from the imager library. Watershed algorithms are frequently used in topograhical analysis to outline drainage basins. Given the morphological similarity between an inverted canopy and a terrain model, this same process can be used to outline tree crowns. However, a potential problem is the issue of oversegmentation, whereby branches, bumps and other spurious treetops are given their own segments. This source of error can be mitigated by using a variant of the algorithm known as marker-controlled segmentation (Beucher & Meyer, 1993), whereby the watershed algorithm is constrained by a set of markers–in this case, treetops.
The mcws function also takes a minHeight argument, although this value should be lower than that which was assigned to vwf. For the latter, minHeight defines the lowest expected treetop, whereas for the former it should correspond to the height above ground of the fringes of the lowest trees.
```{r}
# Create crown map
crowns <- mcws(treetops = ttops, CHM = kootenayCHM, minHeight = 1.5, verbose = FALSE)

# Plot crowns
plot(crowns, col = sample(rainbow(50), length(unique(crowns[])), replace = TRUE), legend = FALSE, xlab = "", ylab = "", xaxt='n', yaxt = 'n')
```
By default, mcws returns a raster, where each crown is given a unique cell value. Depending on the intended purpose of the crown map, it may be preferable to store these outlines as polygons. Setting the format argument to “polygons” will convert the rasterized crown map to a set of polygons (a SpatialPolygonsDataFrame). As an added benefit, these polygons will inherit the attributes of the treetops from which they were generated, such as height. Furthermore, an extra attribute, crownArea, will be calculated for each polygon.
It should be noted, however, that producing crown outlines as polygons requires significantly more processing time and disk space.
```{r}
# Create polygon crown map
crownsPoly <- mcws(treetops = ttops, CHM = kootenayCHM, format = "polygons", minHeight = 1.5, verbose = FALSE)

# Plot CHM
plot(kootenayCHM, xlab = "", ylab = "", xaxt='n', yaxt = 'n')

# Add crown outlines to the plot
plot(crownsPoly, border = "blue", lwd = 0.5, add = TRUE)
```
Assuming that each crown has a roughly circular shape, we can use the crown’s area to compute its average circular diameter.
```{r}
# Compute average crown diameter
crownsPoly[["crownDiameter"]] <- sqrt(crownsPoly[["crownArea"]]/ pi) * 2

# Mean crown diameter
mean(crownsPoly$crownDiameter)
## [1] 2.882985
```




### References
Popescu, S. C., & Wynne, R. H. (2004). Seeing the trees in the forest. Photogrammetric Engineering & Remote Sensing, 70(5), 589-604.
Beucher, S., and Meyer, F. (1993). The morphological approach to segmentation: the watershed transformation. Mathematical morphology in image processing, 433-481.
