Reading and writing geospatial data in R
=============================================
### The raster and sp libraries
The Forest Tools package is built on the raster and sp libraries, which are automatically installed when ForestTools is downloaded. These libraries define a variety of classes and functions for working with raster and vector datasets in R.
It is recommended that any user performing geospatial analyses in R be familiar with both of these libraries. Relatively easy and straightforward guides for raster and sp have been written by their respective authors.

### Geospatial classes used by Forest Tools
* Data product
* Data type
* Object class
* Canopy height model
* Single-layer raster
* RasterLayer
* Treetops
* Points
* SpatialPointsDataFrame
* Crown outlines
* Polygons
* RasterLayer, SpatialPolygonsDataFrame
* Gridded statistics
* Multi-layer raster
* RasterLayer, RasterBrick


### Raster files
To load a raster file, such as a CHM, use the raster function from the ***raster*** library (both the function and the library have the same name). Simply provide a path to a valid raster file. Don’t forget to use either double backslashes \\ or forward slashes / in the file path.

```{r}
library(raster)

# Load a canopy height model
inCHM <- raster("C:\\myFiles\\inputs\\testCHM.tif")
```
Once you have performed your analysis, use the writeRaster function to save any raster files you have produced. 
Setting an appropriate dataType is optional, but can save disk space.
```{r}
# Write a crown map raster file
writeRaster(crowns, "C:\\myFiles\\outputs\\crowns.tif", dataType = "INT2U")
```

### Polygon and point files
There are many options for saving point and polygon files to disk. The ***rgdal*** library provides functions for reading and writing the most common vector formats. The following examples use ESRI Shapefiles.
Use the ```readOGR()`` function to load a polygonal ESRI Shapefile. Instead of providing an entire file path,``readOGR()`` takes two 
separate arguments: the file’s directory, followed by the file name without an extension. 

The following would import a file named ``“C:\myFiles\blockBoundaries\block375.shp”``.
```{r}
library(rgdal)

# Load the 'block375.shp' file
blk375boundary <- readOGR("C:\\myFiles\\blockBoundaries", "block375")
```
Follow this same convention for saving a vector file to disk using writeOGR. A driver must also be specified.
```{r}
# Save a set of dominant treetops
writeOGR(ttops, "C:\\myFiles\\outputs", "treetops", driver = "ESRI Shapefile")
```
